### 2016-02-03 Projects/DelCo_2012/Report/DelCo_2012.tex
edits to CAF report

>>>>

Water is the most widely used fire extinguishing agent because it is effective, environmentally friendly, nontoxic, inexpensive and in many cases, readily available. In addition, water has a very high heat of vaporization per unit mass, at least four times as high as that of any other nonflammable liquid~\cite{NFPA}. The latent heat of vaporization of water is 2255~kJ/kg (970~Btu/lb.)~\cite{NFPA}. This means that 2255~kJ (2139~Btu) of energy is required to change 1~kg (2.2~lb.) of water into steam.  When water is vaporized, its volume increases approximately 1,600 times. Because the energy absorbing capabilities of water are well quantified, they can be used as a basis to calculate the theoretical minimum delivery rate of water needed to extinguish a burning material with a known heat (energy) release rate. Unfortunately, experience has shown that water must be applied at 10 to 100 times the theoretical rate in practice to control and extinguish the fire~\cite{Friedman:2}. As a result of this apparent inefficiency and the need to address fires containing a wide variety of materials, water-based fire fighting additives have been utilized for many years to enhance the fire fighting capabilities of ordinary water.

The most widely used of the water-based agents are the foams used on Class B (liquid hydrocarbon) fires. Agents designed primarily for Class A (ordinary combustible) fires have been used most extensively in conjunction with wildland fires. More recently these agents have been promoted for use on a wider range of Class A and in some cases Class B and D (combustible metal) fires. These agents are frequently claimed to be more effective than plain water while being environmentally safe. In some cases, they are also claimed to reduce the quantity and toxicity of smoke. To gain an understanding of how additives might enhance the fire suppression capabilities of water, it is important to examine the principles of fire suppression.
